env_and_copies_stats

### 2023-04-03 code/analysis_code/stats/cca.qmd
edits

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---

### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

@@

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

@@

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master0 %>%
  select(copies_mL, t7sum, site, date) 
```

@@

@@

@@

@@

# Blind Oso Site:

### Add one day lag

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

### Plot

```{r}
bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)")
```

### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to  data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo2)

#View as Tibble 
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_bo <- augment(growth_fit_bo, bo2)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

Approaches based on such selection criteria essentially try to guess how the model would perform if it were to be fit to new data, without actually trying to do it (in contrast to CV).

```{r}
bo_mod<- lm(copies_mL ~ lag, data = bo2)
AIC(bo_mod)
```

### Residuals

```{r}
res_bo<- resid(bo_mod)
plot(fitted(bo_mod), res_bo)
abline(0,0)
```

# Canals:

#### Add two day lag

```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

### Plot

```{r}
c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)")
```

### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to  data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c22)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_c2 <- augment(growth_fit_c2, c22)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
c2_mod<- lm(copies_mL ~ lag, data = c22)
AIC(c2_mod)
```

### Residuals

```{r}
res_c2<- resid(c2_mod)
plot(fitted(c2_mod), res_c2)
abline(0,0)
```

## Gulf Site:

### Add two day lag

```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

### Plot

```{r}
rd2 %>% ggplot(aes(lag, copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)")
```

### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to  data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd2)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_rd <- augment(growth_fit_rd, rd2)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
rd_mod<- lm(copies_mL ~ lag, data = rd2)
AIC(rd_mod)
```

### Residuals

```{r}
res_rd<- resid(rd_mod)
plot(fitted(rd_mod), res_rd)
abline(0,0)
```

# TEMPERATURE

### Blind Oso

>>>>

---
title: "cca"
format:
  html:
    theme: default
---

# Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
```

@@

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)

@@

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
master<- master0[c(1:10,15,33)]

```

@@

**Need to include 7/5 and 7/6 for previous dust data (will be used for lag analysis)**

@@

\*\*Dust (t7sum) was log transformed

@@

**For this, we cannot have NAs so we choose the strict daily time series**

\***Remember that dust is log transformed**

@@

+------+-----+
| Site | Lag |
+======+=====+
| BO   | 1   |
+------+-----+
| C2   | 2   |
+------+-----+
| RD   | 2\* |
+------+-----+

: \*[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022. Our copies_mL data only contains values from 7/7/22 --> 7/19/22. Therefore, these CCA were made with only this time series. This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8. But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days. Below we will run linear models that DO contain these prior dust days. And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs. But this ultiamtely means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code. Using a combination of visual and CCA would be helpful.

### 2023-04-03 code/analysis_code/stats/distribution.qmd
edits

    A p-value > 0.05 implies that the distribution of the data is **not significantly differen**t from normal distribution. In other words, **we can assume the normality.**

@@

>>>>

    A p-value > 0.05 implies that the distribution of the data is **not significantly differen**t from normal distribution. In other words, **we can assume the normality for the copies_mL data.**

@@

**Based on the Shapiro test, the dust permutation that we are using (t7sum) is not normally distributed. After log transforming, we see that this is now normally distributed. For downstream analysis, we will need to use the log10(dust).**

### 2023-04-03 code/analysis_code/stats/uni_variate.qmd
edits

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

@@

**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input. This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date. We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**

## Cross-correlation:

```{r}
#Blind Oso
bo_ccf<- bo %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
title("Blind Oso DustxCopies CCA", line =1)

#Canals
c2_ccf<- c2 %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
title("Canals DustxCopies CCA", line=1)

#Gulf
rd_ccf<- rd %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
title("Gulf DustxCopies CCA", line =1)
```

**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1**

# Blind Oso Site:

@@

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

# TEMPERATURE

### Blind Oso

>>>>

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sumlog = log10(t7sum)) #Log transform Dust conc (as it is not normal)

@@

# Blind Oso:

@@

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sumlog, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sumlog, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sumlog, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

@@

# SALINITY

## Blind Oso

#### New Dataframe

```{r}
master2<- 
  master0 %>%
  select(copies_mL,sal,temp,site)
```

#### Plot

```{r}
master2 %>%
  filter(site %in% "Blind Oso") %>%
  ggplot(aes(
    x = sal, 
    y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sal, y = copies_mL),
               rr.digits = 4) +
  labs(title = "Sal x Copies BO")
```
